---
title: "reproducible_research_week2_part1"
author: "tge"
date: "7/21/2018"
output: html_document
---

##Loading and preprocessing the data
```{r read, echo=TRUE}
data <- read.csv("~/Downloads/activity.csv")
```

##What is mean total number of steps per day? 

###1. Calculate the total number of steps taken per day
```{r step-day, echo=TRUE}
#aggregate the data set by summing steps in the same day
data_day <- aggregate(data$steps, by=list(data$date), sum)
#rename columns
names(data_day)[1] = "date"
names(data_day)[2] = "total_steps"
#show head of the new dataset
head(data_day)
```

###2. Make a histogram of the total number of steps taken each day
```{r histogram, echo=TRUE}
#plot total daily steps
library(ggplot2)
ggplot(data_day, aes(x=total_steps)) + 
  geom_histogram(color = "black", binwidth = 1200) + 
  labs(title = "Histogram of Total Daily Steps", x = "Daily Steps", y = "Number of Days")
```

###3. Calculate and report the mean and median of the total number of steps taken per day
```{r mean}
mean(data_day$total_steps, na.rm = TRUE)
```
```{r median}
median(data_day$total_steps, na.rm = TRUE)
```

##What is the average daily activity pattern?

###1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r daily, echo=TRUE}
#subset of the data with NA removed
data_clean <- data[!is.na(data$steps),]
#aggregate the clean data set by taking the mean for each time interval
data_clean_mean <- aggregate(data_clean$steps, by=list(data_clean$interval), mean)
#rename columns of data_clean_mean
names(data_clean_mean)[1] = "interval"
names(data_clean_mean)[2] = "mean_steps"
#time series plot
ggplot(data_clean_mean, aes(x=interval, y=mean_steps))+
  geom_line(color="black")+
  labs(title="Mean Steps vs. Time Interval", x="interval", y="mean steps")
```

###2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max, echo=TRUE}
max_mean <- data_clean_mean[which.max(data_clean_mean$mean_steps),]
max_mean
```
