---
title: "reproducible_research_week2_part2"
author: "tge"
date: "7/21/2018"
output: html_document
---

##Loading and preprocessing the data
```{r read, echo=TRUE}
data <- read.csv("~/Downloads/activity.csv")
```

##Imputing missing values

###1. Calculate and report the total number of missing values in the dataset.
```{r missing, echo=TRUE}
missing <-sum(is.na(data$steps))
missing
```

###2. Devise a strategy for filling in all of the missing values in the dataset.
Strategy is using mean steps for a time interval across all days to replace the missing steps for the same time interval at any day

###3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r filled, echo=TRUE}
library(magrittr)
library(dplyr)
#new dataset is named data2
data2 <- data
#replace the missing value with the mean for the corresponding interval
replace_with_mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
data2 <- data2%>% group_by(interval) %>% mutate(steps=replace_with_mean(steps))
#check no missing value
sum(is.na(data2$steps))
```

###4a. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r histogram_new, echo=TRUE}
library(ggplot2)
#aggregate the new data set by summing steps in the same day
data2_day <- aggregate(data2$steps, by=list(data2$date), sum)
#rename columns
names(data2_day)[1] = "date"
names(data2_day)[2] = "total_steps"
#show head of the new dataset
head(data2_day)
#plot total daily steps for new data
ggplot(data2_day, aes(x=total_steps)) + 
  geom_histogram(fill = rgb(0,0,1,1/4), binwidth = 1200) + 
  labs(title = "Histogram of Total Daily Steps", x = "Daily Steps", y = "Number of Days")
```

###4b. Do these values differ from the estimates from the first part of the assignment? 

```{r step-day, echo=TRUE}
#aggregate the data set by summing steps in the same day
data_day <- aggregate(data$steps, by=list(data$date), sum)
#rename columns
names(data_day)[1] = "date"
names(data_day)[2] = "total_steps"
#show head of the new dataset
head(data_day)
```

```{r compare_mean, echo=TRUE}
old_mean <- mean(data_day$total_steps, na.rm = TRUE)
new_mean <- mean(data2_day$total_steps)
old_mean
new_mean
new_mean - old_mean
```

No Change in the mean.

```{r compare_median, echo=TRUE}
old_median <- median(data_day$total_steps, na.rm = TRUE)
new_median <- median(data2_day$total_steps)
old_median
new_median
new_median - old_median
```

Median increases by 1.19.

###4c. What is the impact of imputing missing data on the estimates of the total daily number of steps?

Move the estimates towards the mean. Counts for mean steps increase.

##Are there differences in activity patterns between weekdays and weekends?

###1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r level, echo=TRUE}
#convert to Data format
data2$weekday <- weekdays(as.Date(data2$date))
#add level factor "Weekend, Weekday"
data2$level <- ifelse(data2$weekday=="Saturday" | data2$weekday=="Sunday", "Weekend", "Weekday" )
```

###2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).  
```{r plot2, echo=TRUE}
#aggregate data2 by level and by interval
data2_mean <- aggregate(data2$steps, by=list(data2$level, data2$interval), mean)
#rename the columns
names(data2_mean)[1]="level"
names(data2_mean)[2]="interval"
names(data2_mean)[3]="mean_steps"
#plot
ggplot(data2_mean, aes(x = interval, y=mean_steps, color=level)) +
  geom_line() +
  facet_grid(level ~ .) +
  labs(title = "Mean Steps for Each Interval", x = "interval", y = "mean steps")
```
